---
title: "Data analysis_BMC Public Health"
author: "Martin Tušl"
date: "12/22/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, error=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Packages used for the analysis
```{r}
library(tidyverse)
library(MASS)
library(car)
```

#### Checking for multicollinearity
```{r}
m <- read.csv("~/Desktop/COVID-19 study/R/Covid-19/data.csv", header=TRUE)
m$Work.wors <- as.numeric(m$Work.wors)
m$Age.cat <- as.factor(m$Age.cat)
m$Contract.change <- as.factor(m$Contract.change)
m$Ovrl_HO <- as.factor(m$Ovrl_HO)
m$Freetime <- as.factor(m$Freetime)
m$Work_time <- as.factor(m$Work_time)
m$Caring_duties <- as.factor(m$Caring_duties)
mod1 <- lm(Work.wors ~ Gender + Country + Age.cat + relevel(Contract.change, ref = "No change") + Liv.situation + relevel(Ovrl_HO, ref = "No HO") + relevel(Freetime, ref = "Unchanged") + relevel(Work_time, ref = "Unchanged") + relevel(Caring_duties, ref = "Unchanged"), data = m)
summary(mod1)
exp(coefficients(mod1))
exp(confint(mod1))

vif(mod1)
```

#### Ordinal regression analysis

### Work life worsened as a dependent variable
```{r}
m$Work.wors <- as.ordered(m$Work.wors)
m$Work.wors <- as.ordered(m$Work.wors)
m$Work.imp <- as.ordered(m$Work.imp)
m$Privat.wors <- as.ordered(m$Privat.wors)
m$Privat.imp <- as.ordered(m$Privat.imp)
ord1 <- polr(Work.wors ~ Gender + Country + Age.cat + relevel(Contract.change, ref = "No change") + Liv.situation + relevel(Ovrl_HO, ref = "No HO") + relevel(Freetime, ref = "Unchanged") + relevel(Work_time, ref = "Unchanged") + relevel(Caring_duties, ref = "Unchanged"), data = m, Hess = TRUE)
summary(ord1)
```

### Calculating p values, OR, and CI 
```{r}
ctable1 <- coef(summary(ord1))
p <- pnorm(abs(ctable1[, "t value"]), lower.tail = F) * 2
ctable1 <- cbind(ctable1, "p value" = p)
ci1 <- confint(ord1)
exp(coef(ord1))
orci1 <- exp(cbind(OR = coef(ord1), ci1))
dat1 <- data.frame(orci1, round(orci1, 2))
dat2 <- data.frame(p,round(p, 4))
```

### Work life improved as a dependent variable
```{r}
ord2 <- polr(Work.imp ~ Gender + Country + Age.cat + relevel(Contract.change, ref = "No change") + Liv.situation + relevel(Ovrl_HO, ref = "No HO") + relevel(Freetime, ref = "Unchanged") + relevel(Work_time, ref = "Unchanged") + relevel(Caring_duties, ref = "Unchanged"), data = m, Hess = TRUE)
summary(ord2)
```

### Calculating p values, OR, and CI 
```{r}
ctable2 <- coef(summary(ord2))
p <- pnorm(abs(ctable2[, "t value"]), lower.tail = F) * 2
ctable2 <- cbind(ctable2, "p value" = p)
ci2 <- confint(ord2)
exp(coef(ord2))
orci2 <- exp(cbind(OR = coef(ord2), ci2))
round(orci1, 2)
dat1 <- data.frame(orci2, round(orci2, 2))
dat2 <- data.frame(p,round(p, 4))
```

### Private life worsened as a dependent variable
```{r}
ord3 <- polr(Privat.wors ~ Gender + Country + Age.cat + relevel(Contract.change, ref = "No change") + Liv.situation + relevel(Ovrl_HO, ref = "No HO") + relevel(Freetime, ref = "Unchanged") + relevel(Work_time, ref = "Unchanged") + relevel(Caring_duties, ref = "Unchanged"), data = m, Hess = TRUE)
summary(ord3)
```

### Calculating p values, OR, and CI 
```{r}
ctable3 <- coef(summary(ord3))
p <- pnorm(abs(ctable3[, "t value"]), lower.tail = F) * 2
ctable2 <- cbind(ctable3, "p value" = p)
ci3 <- confint(ord3)
exp(coef(ord3))
orci3 <- exp(cbind(OR = coef(ord3), ci3))
dat1 <- data.frame(orci3, round(orci3, 2))
dat2 <- data.frame(p,round(p, 4))
```

### Privat life improved as a dependent variable
```{r}
ord4 <- polr(Privat.imp ~ Gender + Country + Age.cat + relevel(Contract.change, ref = "No change") + Liv.situation + relevel(Ovrl_HO, ref = "No HO") + relevel(Freetime, ref = "Unchanged") + relevel(Work_time, ref = "Unchanged") + relevel(Caring_duties, ref = "Unchanged"), data = m, Hess = TRUE)
summary(ord4)
```

### Calculating p values, OR, and CI
```{r}
ctable4 <- coef(summary(ord4))
p <- pnorm(abs(ctable4[, "t value"]), lower.tail = F) * 2
ctable4 <- cbind(ctable4, "p value" = p)
ci4 <- confint(ord4)
exp(coef(ord4))
orci4 <- exp(cbind(OR = coef(ord4), ci4))
round(orci4, 2)
dat1 <- data.frame(orci4, round(orci4, 2))
dat2 <- data.frame(p,round(p, 4))
```

### Mental well-being as a dependent variable
```{r}
m$mwb.cat <- as.factor(m$mwb.cat)
ord5 <- polr(mwb.cat ~ Gender + Country + Age.cat + relevel(Contract.change, ref = "No change") + Liv.situation + relevel(Ovrl_HO, ref = "No HO") + relevel(Freetime, ref = "Unchanged") + relevel(Work_time, ref = "Unchanged") + relevel(Caring_duties, ref = "Unchanged") + Work.wors + Work.imp.3 + Privat.wors.3 + Privat.imp.3, data = m, Hess = TRUE)
summary(ord5)
```

### Calculating p values, OR, and CI
```{r}
ctable5 <- coef(summary(ord5))
p <- pnorm(abs(ctable5[, "t value"]), lower.tail = F) * 2
ctable5 <- cbind(ctable5, "p value" = p)
ci5 <- confint(ord5)
exp(coef(ord5))
orci5 <- exp(cbind(OR = coef(ord5), ci5))
round(orci5, 2)
dat2 <- data.frame(p,round(p, 4))
dat1 <- data.frame(orci5, round(orci5, 2))
```

### Self rated health as a dependent variable
```{r}
m$srh <- as.factor(m$srh)
ord6 <- polr(srh ~ Gender + Country + Age.cat + relevel(Contract.change, ref = "No change") + Liv.situation + relevel(Ovrl_HO, ref = "No HO") + relevel(Freetime, ref = "Unchanged") + relevel(Work_time, ref = "Unchanged") + relevel(Caring_duties, ref = "Unchanged") + Work.wors + Work.imp + Privat.wors + Privat.imp, data = m, Hess = TRUE)
summary(ord6)
```

### Calculating p values, OR, and CI
```{r}
ctable6 <- coef(summary(ord6))
p <- pnorm(abs(ctable6[, "t value"]), lower.tail = F) * 2
ctable6 <- cbind(ctable6, "p value" = p)
ci6 <- confint(ord6)
exp(coef(ord6))
orci6 <- exp(cbind(OR = coef(ord6), ci6))
round(orci6, 2)
dat2 <- data.frame(p,round(p, 4))
dat1 <- data.frame(orci6, round(orci6, 2))
```
